---
title: "Monte Carlo Simulation & Confidence Intervals"
author: "Nina van Gerwen (1860852)"
date: "2022-09-14"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Simulation

The simulation is done by the following code.

```{r simulation}
## set seed
set.seed(1248)
## set number of simulations
nsim <- 100
## create an empty data set for the statistics we want
statistics <- as.data.frame(matrix(data = NA, nrow = nsim, ncol = 4))
## create a for loop where in each loop:
for(i in 1:nsim){
  ## 100 samples are gained from a standard normal distribution
  temp_save <- rnorm(100)
  ## calculate the mean (i.e., the bias) of this sample
  statistics[i, 1] <- mean(temp_save)
  ## calculate the standard error of this sample
  statistics[i, 2] <- sd(temp_save)/sqrt(100)
  ## calculate the lower bound 95% CI
  statistics[i, 3] <- statistics[i, 1] - 1.96 * statistics[i, 2]
  ## calculate the upper bound 95% CI
  statistics[i, 4] <- statistics[i, 1] + 1.96 * statistics[i, 2]
  ## and save all those values in the correct place
}
```

### Confidence interval plot:

```{r plot}
library(tidyverse)
library(ggplot2)

statistics$Group <- ifelse(statistics$V3 > 0 & statistics$V4 > 0 
                           | statistics$V3 < 0 & statistics$V4 < 0, FALSE, TRUE)

ciplot <- ggplot(data = statistics, aes(x = 1:100, y = V1, col = Group)) +
    geom_point() +
  geom_abline(col = "blue", slope = 0, intercept = 0) +
  geom_errorbar(ymin = statistics$V3, ymax = statistics$V4) +
  ylim(-0.5, 0.5)

ciplot
```

### Table for simulated samples that were significantly from 0

Due to how my for loop is coded, I can't get all the simulated samples (they
were not saved to save space). Instead have this very not useful table
that shows the mean, standard error and confidence interval of the
simulated samples that did not contain the population value.

```{r}
library(tableone)

statistics %>%
  subset(., Group == FALSE)
```

