---
title: 'Exercise 2: Workflow in RMarkdown'
author: "Nina van Gerwen (1860852)"
date: "2022-10-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## A Reproducible Exercise in RMarkdown

For my exercise, I have chosen to simulate dichotomous IRT data. This
can be done with the following code.

```{r}
## First, set a seed for reproducibility
set.seed(3)

## Decide on the sample size and test length
n <- 100
k <- 5

## Sample person parameters from a standard normal distirbution
theta <- rnorm(n)

## Create a matrix that repeats theta k times so we can use
## the hadamard product later
theta_k <- matrix(data = rep(theta, k), ncol = k)

## Choose item parameters for every item and put this in a matrix
    ## item difficulty
beta <- matrix(data = rep(c(-2, -1, 0, 1, 2), k), ncol = k, nrow = n,
               byrow = TRUE)
    ## item discrimination
alpha <- matrix(data = 0.7, ncol = k, nrow = n)

## Calculate the probability to score a 1 for every person on every item
## following the 2PL model in IRT
Z <- exp((alpha * theta_k) + beta) / (1 + exp((alpha *theta_k) + beta))

## Create a matrix of dichotomous data that can be seen as answers of
## 100 sampled participants to a dichotomous test with 5 items
data <- matrix(data = rbinom(n = n * k, size = 1, prob = Z), ncol = k,
               nrow = n)

## Check the summary of the data
summary(data)
```

From the summary statistics, we find that item 1 is the most difficult as the 
least people had it correct (13%), whereas item 5 was the easiest with 
86% of the sample scoring a 1.

To replicate these results, we can simply run only the following code. However,
this assumes that the item parameters are identical to the first time (which
should be the case in a replication).

```{r Replicating results}
## Sample new person parameters from a standard normal distribution
rep_theta <- rnorm(n)

## Create a matrix that repeats theta k times so we can use
## the hadamard product later
rep_theta_k <- matrix(data = rep(theta, k), ncol = k)

## Calculate the probability to score a 1 for every person on every item
## following the 2PL model in IRT
Z <- exp((alpha * rep_theta_k) + beta) / (1 + exp((alpha * rep_theta_k) + beta))

## Create a matrix of dichotomous data that can be seen as answers of
## 100 sampled participants to a dichotomous test with 5 items
rep_data <- matrix(data = rbinom(n = n * k, size = 1, prob = Z), ncol = k,
               nrow = n)

## Check the summary of the data
summary(rep_data)
```

This time we find that 15% scored a 1 on the most difficult item (V1) and 90%
scored a 1 on the easiest item (V5). This shows that there is variance between
different studies, which is to be expected as we are working with probabilities.

### Session Information

In order to allow people to know under which settings the above exercise was done,
we have included the session information.

```{r}
sessionInfo()
```


