---
title: "Ex2_KP"
author: "Kyuri Park"
date:  "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE, 
                      comment = NA)
```
<style type="text/css">
@import url('https://fonts.googleapis.com/css2?family=Lato:wght@300;400&display=swap');

body{ /* Normal  */
  font-size: 13px;
  font-family: 'Lato', sans-serif;
  }
h1.title {
  font-size: 25px;
  color: DarkBlue;
  margin-bottom:5px;
}
h3.subtitle{ /* Subtitle */
  font-size: 20px;
  color: DarkBlue;
  margin-top:0;
}
h1 { /* Header 1 */
  font-size: 20px;
  font-weight: bold;
}
h2 { /* Header 2 */
  font-size: 17px;
  line-height: 1.6;
}
h3 { /* Header 3 */
  font-size: 15px;
  line-height: 1.6;
}
pre { /* Code block - determines code spacing between lines */
  font-size: 13px;
}

</style>
<hr>

# Aim

1. We explore how we can derive a linear [Equilibrium Causal Model](https://link.springer.com/article/10.1007/s10670-021-00463-0) (ECM) from a random-intercept VAR(1) model. 
2. Upon deriving the ECM, we simulate data from the corresponding model. 
3. We try recovering the model parameters based on the simulated data.

---

# Set up

## Load packages
```{r}
library(qgraph) # for visualizing network
library(lavaan) # for fitting SEM model
library(MASS) # for generating data
library(future) # for parallel processing
library(dplyr)
library(furrr)
```

# Specify VAR(1) model

VAR(1) model can be written as:
$$ X_t = c + BX_{t-1} + \epsilon_t, \ where \ \epsilon \sim N(0, \Sigma_{\epsilon}) $$
Let's specify the parameters of the VAR(1) model. For the sake of simplicity, we will set the number of variables to four (i.e., `p = 4`).  $B$ represents the lagged effects and the values are specified such that it can reach equilibrium (i.e., all $|eigenvalues| \leq 1$).

```{r}
# number of variable
p = 4
# matrix lagged parameters
B <- matrix(c(.5,0,0,0,
              .3,.4,.2,0,
              0,.2,.2,-.4,
              0,0,0,.4), p, p,byrow = T)

# check stability
all(abs(Re(eigen(B)$values)) <= 1)

# network visualization
layout <- rbind(c(0,1), 
                c(1,1), 
                c(1,0),
                c(0,0))
qgraph(t(B), edge.labels = TRUE, layout = layout, fade = FALSE, labels=LETTERS[1:p])
```

# Derive ECM
We can acquire equilibrium model by re-scaling $B$ matrix (marginalizing out diagonal) ([Hyttinen, 2012](chrome-extension://efaidnbmnnnibpcajpcglclefindmkaj/https://jmlr.org/papers/volume13/hyttinen12a/hyttinen12a.pdf)).
```{r}
# Re-scaling function for the beta matrix
rs_beta <- function(B){
  p <- ncol(B)
  ind <- which(diag(B)!=0)
  if(length(ind)== 0){ return(B)} else {
    U <- matrix(0,p,p)
    # Hytinnen - apply the following repeatedly for all B[i,i] != 0
    for(i in ind){
      # let U be a square 0 matrix with U[i,i] = 1 iff B[i,i] != 0
      U[i,i] <- (B[i,i]/(1-B[i,i]))
    }
    B_t <- B - U%*%(diag(p)-B)
    return(B_t)
  }  
}

# Acquire equilibrium model by re-scaling: marginalizaing out diagonal (AR)
B_eq <- rs_beta(B)
dimnames(B_eq) <- list(LETTERS[1:p], LETTERS[1:p])

# check of stability still holds
all(abs(Re(eigen(B_eq)$values)) <= 1)

# "static/cross-sectional" cyclic causal graph
qgraph(t(B_eq), edge.labels = TRUE, layout = layout, fade = FALSE, labels=LETTERS[1:p])
```

# Extend model (multiple-subject)
Now, we extend our model to the multiple-subject random-intercept case:
$$ X_{i,t} = c_i + BX_{i,t-1} + \epsilon_{t-1} $$ 
We assume the $\epsilon$ has the same variance-covariance matrix across individuals and that $B$ is shared among the subjects. We assume the intercepts ($c_i$) differ across the subjects, hence we specify a variance-covariance matrix for them. For the sake of simplicity, let's say $\mu_c$ (average intercepts across subjects) is 1.

```{r}
# Re-scaling function for the variance covariance matrix of the intercepts
rs_sigma <- function(S, B){
  p <- ncol(S)
  ind <- which(diag(B)!=0)
  if(length(ind)== 0){ return(S)} else {
    S_t <- S
    # Hytinnen - apply the following repeatedly for all B[i,i] != 0
    for(i in ind){
      # let U be a square 0 matrix with U[i,i] = 1 iff B[i,i] != 0
      U <- matrix(0,p,p) ; I <- diag(p)
      U[i,i] <- 1
      S_t <- (I + (B[i,i]/(1-B[i,i]))*U)%*%S_t%*%t(I + (B[i,i]/(1-B[i,i]))*U)
    }
    return(S_t)
  }  
}

# define variance covariance matrix of the intercepts
varint <- 1
Sigma <- diag(varint,p)

# Re-scale covariance matrix for the equilibrium setting
# necessary to generate data correctly under this model
Sigma_eq <- rs_sigma(Sigma, B)
Sigma_eq
```

# Simulate data
We simulate data from the equilibrium model. The steps are as follows:  
1. Fix the random seed.  
2. Generate the random intercepts for each person.  
3. Generate the equilibrium positions by *pushing them through* the matrix with the directed parameters.  

```{r}
# step 1
myseed = 123
nsim = 100 # number of simulations
n <- 50000 # sample size
I <- diag(p) # identity matrix
library(future)
plan(multisession)

sim <- future_map(1:nsim, function(x){
  # step 2
  e_eq <- mvrnorm(n, mu = rep(0,p), Sigma = Sigma_eq)
  # step 3
  mu_data_eq <- as.data.frame(t(sapply(1:n,function(i) solve(I - B_eq)%*%e_eq[i,])))
  # specify column names
  colnames(mu_data_eq)<- LETTERS[1:p]
  
  return(SimData = mu_data_eq)
}, .options = furrr_options(seed = myseed)
)
```

# Fit the model (Recover parameters)

```{r}
# specify model
mod ="
B  ~ A + C
C ~ B + D
"
# fit sem model
fits <- lapply(sim, function(x) sem(mod, data=x, std.ov=FALSE))

# estimates@Model
beta_est <- lapply(fits, function(x) (lavInspect(x, what = "est")$beta[LETTERS[1:p],LETTERS[1:p]]))

# compare with original Beta
beta_est[1:3] ; B_eq
```

# Info session
```{r}
sessionInfo()
```


