---
title: "Exercise 2"
author: "Christoph Völtzke"
date: "2022-10-17"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
This is the file for the second exercise focusing on the following question:

*Design a study that:*\
    - Does something that requires RNG \
    - fixes the RNG seed \
    - Replicates the results \
    - Generates an reproducible archive/reprex/markdown\
    - Will run on my machine without trouble (package installs may be excused)\
    - Communicates the info of your session
    
# Aim

This small study aims to fit classification trees on the widely known titanic data set. As I will use ensemble methods including cross-validation a RNG is needed to assure reproducable results

# Reproducible code

## Fixing the random seed
```{r}
set.seed(123)
```

## Required packages
```{r message=FALSE, warning=FALSE}
library(tidyverse) # Data manipulation
library(magrittr) # Pipes
library(caret)
library(gbm)
library(xgboost)
library(readr) # loading the data
```

## Data set
```{r}
titanic <- read_csv("titanic.csv")
```

### Data manipulation
```{r}
titanic <-  titanic %>% 
  select(Survived,Pclass, Age, Sex) %>% 
  mutate(Survived = as.factor(Survived),
         Age = as.numeric(Age), 
         Sex = as.numeric(factor(Sex)), 
         Pclass = as.numeric(factor(Pclass)))
```

### Separating Training and Test Set
```{r}
N <- nrow(titanic)
idx_train <- sample(1:N, size = round(N * 0.7))
train_df <- titanic[idx_train, ] # Training data
val_df <- titanic[-idx_train, ] # Test data
```

## Validation settings
```{r}
cvcontrol <- trainControl(method = "repeatedcv", 
                          number = 10,
                          allowParallel = TRUE)
```

## Fitting a bagged model
```{r}
bag_train_t <- train(Survived ~ .,
                   data = train_df, 
                   method = 'treebag',
                   trControl = cvcontrol,
                   importance = TRUE)
```

## Fitting a random forest model
```{r}
rf_train_t <- train(Survived ~ .,
                   data = train_df, 
                  method = 'rf',
                  trControl = cvcontrol,
                  importance = TRUE)
```

### Assessing importance of predictors
```{r}
rf_train_t %>%
  varImp %>%
  plot
```

## Fitting a gbm model 
```{r}
gbm_train_t <- train(Survived ~ .,
                   data = train_df,
                   method = "gbm",
                   verbose = F,
                   trControl = cvcontrol)
```

## Fitting a Boosting model with the XGBoost package
```{r}
train_x <- model.matrix(Survived ~ .,
                   data = train_df)[,-1]
train_y <- as.numeric(train_df$Survived) - 1
xgboost_train_t <- xgboost(data = train_x,
                         label = train_y, 
                         max.depth = 6,
                         eta = 0.3,
                         nthread = 4,
                         nrounds = 4,
                         objective = "binary:logistic",
                         verbose = 2)

```

## Assess the predictive performance of the different models
```{r}
bag_test <- predict(bag_train_t, newdata = val_df)
rf_test  <- predict(rf_train_t, newdata = val_df)
gbm_test <- predict(gbm_train_t, newdata = val_df)
xgb_test <- predict(xgboost_train_t, newdata =  model.matrix(Survived ~ .,
                   data = val_df)[,-1]) %>%
  factor(x = ifelse(. < 0.5, 1, 2), levels = c(1,2), labels = c("Survived", "Died"))

table(bag_test, val_df$Survived)
table(rf_test, val_df$Survived)
table(gbm_test, val_df$Survived)
table(xgb_test, val_df$Survived)
```
# Replication
```{r}
sessionInfo()
```

