---
title: "Exercise 1_KP"
author: "Kyuri Park"
date: "Sep 15 2022"
output: 
  html_document:
    toc: false
    toc_depth: 5
    toc_float: true
    number_sections: false
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE, 
                      comment = NA)
```

<style type="text/css">
@import url('https://fonts.googleapis.com/css2?family=Lato:wght@300;400&display=swap');

body{ /* Normal  */
  font-size: 13px;
  font-family: 'Lato', sans-serif;
  }
h1.title {
  font-size: 25px;
  color: DarkBlue;
  margin-bottom:5px;
}
h3.subtitle{ /* Subtitle */
  font-size: 20px;
  color: DarkBlue;
  margin-top:0;
}
h1 { /* Header 1 */
  font-size: 20px;
  font-weight: bold;
}
h2 { /* Header 2 */
  font-size: 17px;
  line-height: 1.6;
}
h3 { /* Header 3 */
  font-size: 15px;
  line-height: 1.6;
}
pre { /* Code block - determines code spacing between lines */
  font-size: 13px;
}

</style>
<hr>


We load the necessary packages
```{r message=FALSE}
library(future)
library(furrr)
library(dplyr)
library(ggplot2)
```

## 1. Perform a small simulation that does the following:
- Sample 100 samples from a standard normal distribution.
- For each of these samples, calculate the following statistics for the mean.

Here, I referred to Gerko's answer using `future` package.
```{r simulation}
# sample 100 samples of size 1000 from std normal dist.
nsim <- 100
plan(multisession)

simulation <- future_map(1:nsim, 
  function(x){
  x <- rnorm(1000, 0, 1)
  m <- mean(x)
  stderr <- sd(x) / sqrt(length(x))
  error <- qnorm(0.975)*stderr
  coverage <- m - error < 0 & m + error > 0
  return(c(Mean = m, AbsBias = mean(x) - 0 , StdErr = stderr, Lower95 = m - error, Upper95 = m + error, Coverage = coverage))
},
# set the seed 
.options = furrr_options(seed = 123)) %>% 
  do.call('rbind', args = .) %>% 
  as_tibble
```

```{r, collapse=TRUE}
# compute the mean of the statistics
simulation %>% colMeans

## Out of 1000 samples, 930 samples contain the popoulation mean.

```

## 2. Create a plot that demonstrates the following:
> “A replication of the procedure that generates a 95% confidence interval that is centered around the sample mean would cover the population value at least 95 out of 100 times” (Neyman, 1934)

```{r plot, fig.width=10}
simulation %>% ggplot(aes(y=Mean, x=1:100, colour = as.factor(Coverage))) + 
  geom_point()+
  geom_errorbar(aes(ymax = Upper95, ymin = Lower95)) +
  geom_hline(aes(yintercept = 0), color = "dark grey", size = 2) + 
  xlab("Simulations") +
  ylab("Means and 95% Confidence Intervals")

```

## 3. Present a table containing all simulated samples for which the resulting confidence interval does not contain the population value.
```{r}
simulation %>% 
  filter(!Coverage) %>% 
  round(3) %>% 
  DT::datatable(options = list(dom = 'tp'))

knitr::knit_exit()
```
