---
title: "Spectator Attendance: Impact COVID 19"
author: "Christoph Völtzke"
date: "2022-11-26"
output: 
  ioslides_presentation:
    logo: spectator_logo_pexels.jpg
bibliography: dshs.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(DiagrammeR)
library(DT)
library(readr)
library(plotly)
library(lme4)
```

```{r, include=FALSE}
dat <- read_csv("D1 (1).csv")
plot_data <- read_csv("plot_data.csv")
dat_new <- read_csv("dat_new.csv")
```
## Expected effect

<center>
```{r, echo=FALSE}
grViz("
  digraph {
    Home_advantage -> {Spectators Home_Team_Points} 
    Spectators -> {Home_Team_Points}
  }")
```
</center>

- Now we want to investigate whether the absence of Spectators while the COVID 19 pandemic leads to a decrease in Home team points.

## Replication & Extension
<div class="columns-2">

- @wunderlich2021does to be replicated:
  - 10 different leagues spanning 6 countries
  - All seasons from 2010/2011 to 2019/2020 
  - Outcome variables differences in points, goals, shots
  
- 
  - Original gets extended with:
  - Recent 2020/2021 and 2021/2022 seasons 
  - Modeling decisions 
  - Exploratory analysis based on: @jimenez2021home & @scoppa2021social
</div>
## Data presentation

```{r, echo=FALSE}
datatable(dat, options = list(pageLength = 5))
```

- Here is an examplary data set retrieved from:
<center>
[http://www.football-data.co.uk](http://www.football-data.co.uk)
</center>

## How points changed over time

```{r, include=FALSE}
# oterwhise the plot collapses
plot_data$Name <- as.character(plot_data$Name)
plot_data[plot_data == "10_Y"] <- "10"
plot_data[plot_data == "10_N"] <- "10.5"
plot_data$Name <- as.numeric(plot_data$Name)
```

```{r, echo=FALSE, warning=FALSE}
p2 <- ggplot(data = plot_data, aes(x=Name, y=Points, group=1)) +
  geom_line(color="black") +
  geom_point(size=2, color="blue") + 
  ylab("Points") + xlab("Season") +
  geom_hline(yintercept=0, linetype="dashed", color = "red") +
  theme_minimal()

ggplotly(p2)
```
## Model used for analysis

\[\begin{align}
  Points_{ijk} = \gamma_{000} + (\gamma_{100}+ u_{10k} + u_{1jk}) * Covid19_{ijk} \\
    + (\gamma_{200}+ \upsilon_{20k} + u_{2jk}) * Season_{ijk} \\
    + \epsilon_{ijk} + \upsilon_{0jk} + u_{00k}
\end{align}\]

- This is the multi level equation including three levels. The main variable of interest is Covid-19
- If this is significant with a negative coefficent there is a significant decrease in points for the home team

## Code for analysis

```{r, eval=FALSE}
model <- lmer(pointsDiff ~ Covid19 + Season_c + (1|Division/Home),
              REML = FALSE, data = dat_new)
```
- This is the multi level model using the lme4 package.

## Results of analysis

```{r, warning=FALSE, echo=FALSE}
model <- lmer(pointsDiff ~ Covid19 + Season_c + (1|Division/Home),
              REML = FALSE, data = dat_new)
summary(model)
```


## References
