---
title: "Solution to Exercise 1 - using `futures`"
author: "Gerko Vink"
date: "Markup Languages and Reproducible Programming in Statistics"
output: 
  html_document:
    toc: false
    toc_depth: 5
    toc_float: true
    number_sections: true
---

<style type="text/css">

body{ /* Normal  */
      font-size: 12px;
  }
td {  /* Table  */
  font-size: 12px;
}
h1.title {
  font-size: 18px;
  color: DarkBlue;
}
h1 { /* Header 1 */
  font-size: 18px;
}
h2 { /* Header 2 */
    font-size: 18px;
}
h3 { /* Header 3 */
  font-size: 18px;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>

---

This is an exercise in Monte Carlo simulation, that also details a reproducible workflow. But we'll consider more about reproducibility in the next meeting. 

---

We load the necessary packages
```{r message=FALSE}
library(future)
library(furrr)
library(dplyr)
library(magrittr)
```

Next, we define the simulation as a multisession `future`. We need to fix the seed here as the parallel sequence should be replicable. 
```{r}
nsim = 100
plan(multisession)
# start future mapping
SIM <- future_map(1:nsim, function(x){
  x <- rnorm(5000, mean = 0, sd = 1)
  M <- mean(x)
  DF <- length(x) - 1
  SE <- 1 / sqrt(length(x))
  INT <- qt(.975, DF) * SE
  return(c(mean = M, 
           bias = M - 0, 
           std.err = SE, 
           lower = M - INT, 
           upper = M + INT, 
           cov = M - INT < 0 & 0 < M + INT))
}, 
.options = furrr_options(seed = 123), 
.progress = TRUE) %>% 
  do.call("rbind", args = .) %>% 
  as_tibble
```

```{r}
SIM %>% colMeans
```

We can see that 96 out of the 100 samples cover the population value. 

To identify the samples for which the population value is not covered, we can use column `cov` as it is already a logical evaluation. 
```{r}
SIM %>% filter(!cov)
```

To present this info as a table, package `DT` is a wonderful extension to use with `rmarkdown` 
```{r, warning=FALSE, message=FALSE}
library(DT)
SIM %>% 
  round(4) %>% 
  datatable()
```

To create a graph that would serve the purpose of the exercise, one could think about the following graph:

```{r warning=FALSE, message=FALSE}
library(ggplot2)
limits <- aes(ymax = SIM$upper, ymin = SIM$lower)
SIM %>% mutate(covered = as.factor(cov)) %>% 
ggplot(aes(y=mean, x=1:100, colour = covered)) + 
  geom_hline(aes(yintercept = 0), color = "dark grey", size = 2) + 
  geom_pointrange(limits) + 
  xlab("Simulations 1-100") +
  ylab("Means and 95% Confidence Intervals")
```

---

End of solution.