---
title: "Markup: Assignment 1"
author: "Shannon Dickson"
date: "`r format(Sys.Date(), '%B %d %Y')`"
output: 
   bookdown::html_document2:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: false
    theme: paper
    
---
<style type="text/css">
  
body{ /* Normal  */
  font-size: 12px;
  }
td {  /* Table  */
  font-size: 12px;
}
h1.title {
  font-size: 18px;
  color: DarkBlue;
}
h1 { /* Header 1 */
  font-size: 18px;
}
h2 { /* Header 2 */
  font-size: 18px;
}
h3 { /* Header 3 */
  font-size: 18px;
}
code.r{ /* Code block */
  font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
  font-size: 14px;
}
</style>

---

```{r setup, echo = FALSE}
library(knitr)
knitr::opts_chunk$set(include = TRUE, 
                      echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
```

# Preparation

The following packages are required.

```{r}
library(dplyr)
library(magrittr)
library(ggplot2)
library(kableExtra)
```

# Simulation

Below I perform a small simulation that samples 100 times from a normal distribution. Sample statistics are collected for each sample; mean, standard error, bias, and the lower and upper confidence intervals. 

```{r}
set.seed(6368683)
nsim <- 100
samples <- data.frame(matrix(nrow = nsim, ncol = 7))
colnames(samples) <- c("iter", "Mean", "Std.Error", "Bias", "LowerCI", "UpperCI", "Coverage")

for (i in 1:nsim){
  x <- rnorm(5000, mean = 0, sd = 1)
  m <- mean(x)
  df <- length(x) - 1
  se <- 1 / sqrt(length(x))
  ci <- qt(.975, df) * se
  
  samples[i, 1] <- i
  samples[i, 2] <- m
  samples[i, 3] <- se
  samples[i, 4] <- m - 0
  samples[i, 5] <- m - ci
  samples[i, 6] <- m + ci
  samples[i, 7] <- (m - ci < 0 & 0 < m + ci)
} %>% 
  as_tibble()

samples$Coverage <- as.integer(samples$Coverage)
```

In our sample the population value covered 95% of the time. 

```{r}
samples %>%
  colMeans %>% 
  kable(digits = 4, format = "html", caption = "Table 1. Summary of the samples") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = F)
```

To identify samples for which the population value is not covered, we need to filter our data by `coverage`. 

```{r}
samples %>% 
  filter(Coverage == 0) %>% 
  kable(digits = 4, format = "html", caption = " Table 2. Samples outside the population value") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = F)
```

## Visualising the samples

Below 100 samples and their 95% confidence intervals are plotted. The samples not covered are in coral. 

```{r}
samples %>% 
  mutate(coverage = as.factor(Coverage)) %>% 
  ggplot(aes(y=Mean,
             x=1:100,
             colour = coverage)) + 
  geom_point() +
  geom_errorbar(aes(ymax = UpperCI, ymin = LowerCI)) +
  geom_hline(aes(yintercept = 0),
             color = "cornsilk3", size = 1) + 
  xlab("N Sim = 100") +
  ylab("Means and 95% CI") +
  scale_colour_manual(values = c("0" = "lightcoral",
                                 "1" = "darkseagreen")) +
  theme_bw()
```

## Results table

Below the samples are presented in a table. You can scroll to view the 100 samples. 

```{r}
samples %>% 
kable(digits = 4, format = "html", caption = "Table 3. Simulated samples (N = 100) and sample statistics") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = T) %>% 
  scroll_box(width = "100%", height = "200px")
```


