---
title: "Ferrets"
author: "Gaja"
output: 
  ioslides_presentation:
    logo: logo.png
    css: logo.css
bibliography: refer.bib
---

## Some info and a fake citation

::: columns-2
-   The ferret (Mustela furo) is a small, domesticated species belonging to the family Mustelidae. The ferret is most likely a domesticated form of the wild European polecat (Mustela putorius), evidenced by their interfertility. Other mustelids include the stoat, badger and mink.

-   Physically, ferrets resemble other mustelids because of their long, slender bodies. Including their tail, the average length of a ferret is about 50 cm (20 in); they weigh between 0.7 and 2.0 kg (1.5 and 4.4 lb); and their fur can be black, brown, white, or a mixture of those colours. In this sexually dimorphic species, males are considerably larger than females (@lkj).
:::

## Two columns? try 2

<div style="float: left; width: 50%;">
The ferret (Mustela furo) is a small, domesticated species belonging to the family Mustelidae. The ferret is most likely a domesticated form of the wild European polecat (Mustela putorius), evidenced by their interfertility. Other mustelids include the stoat, badger and mink.
</div>

<div style="float: right; width: 50%;">
Physically, ferrets resemble other mustelids because of their long, slender bodies. Including their tail, the average length of a ferret is about 50 cm (20 in); they weigh between 0.7 and 2.0 kg (1.5 and 4.4 lb); and their fur can be black, brown, white, or a mixture of those colours. In this sexually dimorphic species, males are considerably larger than females (@lkj).
</div>

## A cute image {.flexbox .vcenter}

<img src=ferret.jpg
     width="600" 
     height="500" />
     
## A moving cute image  {.flexbox .vcenter}

<img src=ferret.gif
     width="600" 
     height="500" />
     
## Some code cached, executed or else


```{r, echo=FALSE, cache=TRUE, warning=FALSE}
data(iris)
require(pander)
pander(paste0("specie = ", unique(iris$Species)), caption="Species in 
       iris dataset.")
```

```{r, echo=TRUE, eval=FALSE}
a = 1+1
```

## Some aligning of equations

$$
\begin{align}
1 +1 &= 2 \\
a + b &= c \\
\text{something} + \text{something} &= \text{2 times something} \\
\end{align}
$$

## Some datatabling

```{r, echo=FALSE, warning=FALSE}
require(DT)
datatable(iris, options = list(pageLength = 5))
```


## Literature