---
title: "Markdown Exercise 1"
author: "Lela Roos (1729683)"
date: "9/16/2022"
output: html_document
---

# Monte Carlo simulation exercise
```{r}
library(ggplot2)
library(dplyr)
```
```{r}
set.seed(563)
```


## Sample 100 samples

```{r}
n <- 600 # set sample size
samps <- matrix(nrow = 100, ncol = n)

for (i in 1:100){
  samp_i <- rnorm(n)
  for (j in 1:n){
    samps[i,j] <- samp_i[j]
  }
}


```

## Calculations
For each of these samples, calculate the following statistics for the mean:
standard error
lower bound of the 95% confidence interval
upper bound of the 95% confidence interval

```{r}
stats <- function(samp){
  bias <- mean(samp)
  se <- sd(samp)/sqrt(length(samp))
  CI_l <- mean(samp) - 1.96 * se
  CI_u <- mean(samp) + 1.96 * se
  results <- cbind(bias, se, CI_l, CI_u)
  return(results)
}
  
summstats <- matrix(nrow = 100, ncol = 4)

for (i in 1:100){
  summs <- stats(samps[i,])
  for (j in 1:4){
    summstats[i,j] <- summs[j]
  }
}
```


```{r}
summstats <- as.data.frame(summstats)
colnames(summstats) <- c("Bias", "SE", 
                         "Lowerbound_CI", "Upperbound_CI")
summstats <- summstats  %>% 
  mutate(id = row_number(), .before = 1)

samps <- as.data.frame(samps)
samps <- samps  %>% 
  mutate(id = row_number(), .before = V1)

```

## Plot
"A replication of the procedure that generates a 95% confidence interval that is centered around the sample mean would cover the population value at least 95 out of 100 times" (Neyman, 1934)

```{r}
ggplot(summstats, 
       aes(x = 1:nrow(summstats),
           y = Bias)) +
  labs(x="Mean", y = "Index") +
  ylim (-0.3, 0.3) +
  geom_errorbar(aes(ymin=Lowerbound_CI, ymax=Upperbound_CI)) +
  geom_point() + 
  geom_hline(yintercept=0, linetype="dashed", color = "red")

```
## CIs not containing population mean

```{r}
outside_samp_stats <- summstats %>% 
  rename(Mean = Bias) %>% 
  filter(Lowerbound_CI > 0 | Upperbound_CI < 0)

print(outside_samp_stats[,2:5])
```

```{r}
# save samples
outside_samps <- samps %>% 
  filter(id %in% outside_samp_stats$id)
```


