---
title: "Monte Carlo Exercise"
author: "Lotte Mensink"
date: "9/26/2022"
output: html_document
---

First, we sample 100 samples from a standard normal distribution. It is important to set a seed before drawing random samples. 

```{r}
set.seed(123)
nsamp <- 100
samples <- list(nsamp)

for (i in 1:nsamp){
  s <- rnorm(1000)
  samples[[i]] <- s
}
```

For each of these samples, calculate the mean, the absolute bias, the standard error and the 95% confidence interval. 

```{r}
means <- c()
bias <- c()
se <- c()
ll <- c()  
ul <- c()
  
for (i in 1:nsamp){
  data <- samples[[i]]
  means[i] <- mean(data)
  bias[i] <- abs(mean(data) - 0)
  se[i] <- 1/sqrt(length(data))
  int <- qt(0.975, length(data)-1) * se[i]
  ll[i] <- mean(data) - int
  ul[i] <- mean(data) + int
}

results <- as.data.frame(cbind(means, bias, se, ll, ul))

colMeans(results)
```

Create a plot that demonstrates the following:      
*“A replication of the procedure that generates a 95% confidence interval that is centered around the sample mean would cover the population value at least 95 out of 100 times” (Neyman, 1934)*

```{r, message = FALSE}
library(dplyr)
results <- results %>%
  mutate(Covered = ll < 0 & ul > 0)

library(ggplot2)
limits <- aes(ymax = ul, ymin = ll)
ggplot(results, aes(x = 1:100, y = means, color = Covered)) +
    geom_pointrange(limits) + 
    xlab("Simulations") +
    ylab("95% Confidence Intervals")
```

Present a table containing all simulated samples for which the resulting confidence interval does not contain the population value.

```{r}
results %>%
  filter(Covered == FALSE)
```

